\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-05-01 trunk/results.tex
updates

A significant driver of the applicability of our work was the growing
need for AUV missions to be driven by sensor data. This trend is
illustrated by the implementation of MBARI's \emph{Gulper}
water-samplers \cite{Bird07}, $2$ litre syringe style containers $10$
of which are encased within the mid-body section of our Dorado vehicle
as shown in Fig \ref{fig:gulper}. However, the original design and
intended use of operation continued to be dominated by a naive belief
of collecting and returning water samples as traditional ship-borne
methods often employ, namely using pre-scripted methods. With the
installation of \rx onboard the Dorado with its capability sense the
environment and to appropriately synthesize plans and replan on the
fly, more imaginative methods of sample acquisition were targeted.
\kcomment{The Dorado uses two onboard computers: a main vehicle computer which
is a 244 Mhz PC/104 stack running the QNX real-time operating system,
and a 367 MHz EPIC EPX-GX500 AMD Geode stack running Linux and \rxe. A
socket-based communication protocol allows the exchange of goals and
state updates between \rx and the lower level functional layer.}

@@

While the early experiments in \cite{fox2007} coupled the classifier
with the onboard planner, two further novel extensions to sampling
were added. First, to enhance the classifier model by adding a
temporal dimension with a Hidden Markov Model (HMM) \cite{rabiner86,
  Rabiner89atutorial}.  An HMM is a sequential state transition model
in which the current state is not directly visible, but an
observation, dependent on the state, is.  Each state has a probability
distribution over the possible observations.  Therefore the sequence
of observations gives some information about the sequence of traversed
states in the HMM. Unlike identification models based on the simple
classification scheme, HMMs use the dynamics of transition between
modeled states to exploit sequential information of the sensor
readings providing more robust formulation for environmental state
identification. The temporal probabilistic transition then acts as a
filter on top of the clustering model since it is not only derived
from it, but it informs the planner of a potential trigger state only
when the models' transition occurs to an appropriate state of
recognition of the INL feature \cite{mcgann08b}. Not only was
deliberation aided by direct perception of the state (of the INL) but
temporal transition functions added a further level of validation
between expected states in the INL model. This coupling of Machine
Learned model with a deliberative planner is novel in the ocean
sciences and is a continuing direction of investigation
\cite{kumar11}.

@@

To augment our understanding of such microbial communities, the Dorado
with \rx has been utilized in a range of activities in a
quasi-lagrangian mode. As the water mass is being tracked with a
drifting instrument, the vehicle is expected to move around it. While
our model \kcomment{supports} multiples patterns (square, lawn-mower,
transect(s)\dots) with any possible dimension factor, the majority of
at-sea deployments have used a square pattern ($3X3$
km\textsuperscript{2} or a $1X1$ km\textsuperscript{2})\footnote{Often
  post-hoc reconstruction of the data set for visualization drives how
  much adaptation the vehicle can be allowed to undertake.}. Plan
synthesis occurs frequently, but within the context of a drifting
point generated by a drifter. Fig. \ref{fig:dta-setup} shows the
general setup of our experiments which can be tailored with the use of
various computational methods within the \texttt{DTA} app which is
contained within the Oceanographic Decision Support System (\od; see
below) \cite{das11}. GPS data from the free-floating drifter is sent
via satellite and processed onshore to determine a projection (speed
and bearing) of the drifter. This data is sent to the AUV; at the
completion of the current survey, the vehicle receives this data as a
high-level goal which is decomposed and synthesized as a survey
plan. Currently, the vehicle completes the survey without any further
dynamic updates from the drifter. In order to minimize the range of
errors that the survey needs to encapsulate (data latency or surface
or sub-surface currents for those drifters which have a drogue) the
estimation is a projection in time and space of the drifters centroid
sent with this goal from the shore-side app.

\begin{figure}[htpb]
\centering
\includegraphics[width=0.35\textwidth]{figs/june11-drifter-follow-180-181.pdf}
\caption{\small{A June 2011 drifter cruise shown in the Earth
    frame. Deviations in the AUV transect are based on the frame of
    reference in addition to positional errors \cite{das11b}.}}
\label{fig:drifter-errors}
\end{figure}

@@

\begin{figure*}
\centering
\subfloat[]{\label{fig:sep10-ship}\includegraphics[width=0.38\textwidth]{figs/sep10-canon-cruise.png}}
\subfloat[]{\label{fig:sep10-transects}\includegraphics[width=0.35\textwidth]{figs/september-iterations.pdf}}
  \caption{\small{The offshore \can ESP-drifter following cruise
      \cite{das11b}. Fig. \ref{fig:sep10-ship} shows the ESP drifter
      foreground with the R/V \emph{Western Flyer} with the image
      taken from the R/V \emph{Zephyr}. Fig \ref{fig:sep10-transects}
      shows the 86 km transect of the Dorado following the ESP drifter
      about 100 Nm from California shores.}}
\label{fig:gulper}
\end{figure*}

These theoretic results were validated in a five-day long off-shore
deployment carried out in September 2010, 160 kms off the California
coast (Fig.~\ref{fig:sep10-ship}). A specialized drifter with a
genomic sensor \cite{scholin09} hanging $20$ m below, performing
in-situ identification of micro-organisms was the target. The
experiment was supported by crews on two support vessels, the R/V
\emph{Western Flyer} and the R/V \emph{Zephyr}. The \emph{Flyer}
visited the drifter every four hours to carry out a series of
ship-based sampling experiments and lab analysis on water samples to
ground-truth the drifter sensor data. The \emph{Zephyr} was meanwhile
focused on Lagrangian observation studies using the Dorado with the
goal to monitor the nutrient budget at the perimeter of a $1$km $X
1$km water patch around the advecting drifter while the AUV performed
a transformed box pattern around the drifter. A number of logistical
issues were kept in mind while designing and executing the
experiment. Each iteration began with the latest drifter update
(position and velocity) received from the drifter through a satellite
link which was transmitted to the vehicle for in-situ adaptation. \rx
used this input to compute the $5$ waypoints necessary for an
iteration of the box pattern, with the AUV traveling at constant
velocity in the earth frame. Waypoints were computed once at the
beginning of the survey with the AUV surfacing once for every survey,
with each survey lasting $\sim1$ to $1.5$ hours. A total of 60 surveys
were attempted over the course of the $5$
days. Fig. \ref{fig:sep10-transects} shows the overall transects over
the course of the mission, with gaps in between days when the AUV was
being recharged. Each deployment lasted $\sim12$ hours. The black dots
in the figure show the beginning of each iteration when drifter
updates were received by the AUV.

\begin{figure}[t]
\centering
\includegraphics[width=0.7\textwidth]{figs/oct10-ballet.pdf}
\caption{\small An October 2010 robot ``ballet'' where a smaller
  slower AUV tracked a patch center, sending periodic updates to the
  Dorado using a setup shown in Fig. \ref{fig:dta-setup}. The Google
  Earth post-hoc image shows {\color{red}red} track lines of the \rx
  commanded Dorado with the {\color{yellow}yellow} transects of the
  LRAUV. {\color{yellow}Yellow} dots show the track line of a
  GPS-enabled drifter used for seeding the experiment over the
  Monterey Canyon with Santa Cruz, California to the right of the
  figure.}
\label{fig:ballet}
\end{figure}

@@

\begin{figure*}
\centering
\subfloat[]{\label{fig:odss-mtg}\includegraphics[width=0.3\textwidth]{figs/october-canon-meeting.png}}
\subfloat[]{\label{fig:odss-virtual}\includegraphics[width=0.33\textwidth]{figs/virtual-drifter-tool.png}}
\subfloat[]{\label{fig:odss-ipad}\includegraphics[width=0.3\textwidth]{figs/photo-ipad-2.jpg}}
\caption{\small Fig. \ref{fig:odss-mtg} shows a 2010 \can meeting with
  the \od in use. Fig. \ref{fig:odss-virtual} shows a drifter tracking
  tool used during \can planning; it used a physical model of advected
  surface currents from \texttt{CeNCOOS} \cite{cencoos} to plan AUV
  deployments. Fig. \ref{fig:odss-ipad} shows \od use on an iPad.}
\label{fig:odss2}
\end{figure*}

@@

Operationally, the \od has become a software system that the
scientist consults during a \can field campaign. S/he consults
up-to-date model outputs, synoptic views of ocean color from remote
sensing data and weather conditions for the next $24$-hour planning
cycle and then queries the system to compute candidate sampling
locations. In the near future, the \od will consult its situational
awareness database, incorporate model outputs projecting these outputs
into the space-time region of interest to generates candidate
hypotheses with associated utility metrics. With the scientist's
guidance the automated planner would then determine the most viable
robotic option(s) to be dispatched (underwater, surface or aerial),
given available resources, distance of travel, required payload and
the environment. For instance, unpowered autonomous gliders would be
chosen for longer-duration missions with smaller payload requirements
in areas of moderate to weak ocean currents. The automated planner
would balance these criteria to compute a viable trajectory plan for
the selected vehicle. Vehicles with more controllability would be
given a specified area to survey and the path generated \situ. Goals
are communicated to the robot after analysis on shore for selective
targeting of features by the conglomerate of vehicles in the ocean
being tracked by the \od. Thus far, the \od has provided a simple
yet functional tool for scientists to use in the daily complex
decision making and the following experiment and survey design
needs. It has helped train scientists to think of expanding their
repertoire of tools for the scientific process and in the process it
enabled a mixed team of engineers and scientists to interact at the
functional level of the systems' capabilities.

>>>>

A significant driver of the applicability of our work was the growing
need for AUV missions to be driven by sensor data. This trend is
illustrated by the implementation of MBARI's \emph{Gulper}
water-samplers \cite{Bird07}, $2$ litre syringe style containers $10$
of which are encased within the mid-body section of our Dorado vehicle
as shown in Fig \ref{fig:gulper}. However, the original design and
intended use of operation continued to be dominated by a naive belief
of collecting and returning water samples as traditional ship-borne
methods often employ, namely using pre-scripted methods. With the
installation of \rx onboard the Dorado with its capability sense the
environment and to appropriately synthesize plans and replan on the
fly, more imaginative methods of sample acquisition were targeted.
\kcomment{The Dorado uses two onboard computers: a main vehicle
  computer which is a 244 Mhz PC/104 stack running the QNX real-time
  operating system, and a 367 MHz EPIC EPX-GX500 AMD Geode stack
  running Linux and \rxe. A socket-based communication protocol allows
  the exchange of goals and state updates between \rx and the lower
  level functional layer. Event logs from field experiments show that
  a peak load of up to $20\%$ occurs during re-planning. Other desktop
  simulations \cite{mcgann09} show that partitioning scales with
  problem size.  Reactor design and the number of partitions however
  are still problem specific and determined empirically.}

@@

While the early experiments in \cite{fox2007} coupled the classifier
with the onboard planner, two further novel extensions to sampling
were added. First, to enhance the classifier model by adding a
temporal dimension with a Hidden Markov Model (HMM) \cite{rabiner86,
  Rabiner89atutorial}.  An HMM is a sequential state transition model
in which the current state is not directly visible, but an
observation, dependent on the state, is.  Each state has a probability
distribution over the possible observations.  Therefore the sequence
of observations gives some information about the sequence of traversed
states in the HMM. Unlike identification models based on the simple
classification scheme, HMMs use the dynamics of transition between
modeled states to exploit sequential information of the sensor
readings providing more robust formulation for environmental state
identification. The temporal probabilistic transition then acts as a
filter on top of the clustering model \kcomment{from \cite{fox2007},}
since it is not only derived from it, but it informs the planner of a
potential trigger state only when the models' transition occurs to an
appropriate state of recognition of the INL feature
\cite{mcgann08b}. Not only was deliberation aided by direct perception
of the state (of the INL) but temporal transition functions added a
further level of validation between expected states in the INL
model. This coupling of Machine Learned model with a deliberative
planner is novel in the ocean sciences and is a continuing direction
of investigation \cite{kumar11}.

@@

To augment our understanding of such microbial communities, the Dorado
with \rx has been utilized in a range of activities in a
quasi-lagrangian mode. As the water mass is being tracked with a
drifting instrument, the vehicle is expected to move around it. While
our model \kcomment{supports} multiples patterns (square, lawn-mower,
transect(s)\dots) with any possible dimension factor, the majority of
at-sea deployments have used a square pattern ($3X3$
km\textsuperscript{2} or a $1X1$ km\textsuperscript{2})\footnote{Often
  post-hoc reconstruction of the data set for visualization drives how
  much adaptation the vehicle can be allowed to undertake.}. Plan
synthesis occurs frequently, but within the context of a drifting
point generated by a drifter. \kcomment{A drifter tracking application
  (\texttt{DTA}) is used on-shore with an Oceanographic Decision
  Support System (\ode; see below) \cite{das11}, to process the
  drifter location from its onboard GPS and used as a target for the
  Dorado vehicle to provide a configurable and dynamic track-line
  enclosure to provide oceanographic context to the drifting body.}
Fig. \ref{fig:dta-setup} shows the general setup of our experiments
which can be tailored with the use of various computational methods
within the \texttt{DTA} app. GPS data from the free-floating drifter
is sent via satellite and processed onshore to determine a projection
(speed and bearing) of the drifter. This data is sent to the AUV; at
the completion of the current survey, the vehicle receives this data
as a high-level goal which is decomposed and synthesized as a survey
plan. Currently, the vehicle completes the survey without any further
dynamic updates from the drifter. In order to minimize the range of
errors that the survey needs to encapsulate (data latency or surface
or sub-surface currents for those drifters which have a drogue) the
estimation is a projection in time and space of the drifters centroid
sent with this goal from the shore-side app.

\begin{figure}[htpb]
\centering
\includegraphics[width=0.45\textwidth]{figs/june11-drifter-follow-180-181.pdf}
\caption{\small{A June 2011 drifter cruise shown in the Earth
    frame. Deviations in the AUV transect are based on the frame of
    reference in addition to positional errors \cite{das11b}.}}
\label{fig:drifter-errors}
\end{figure}

@@

\begin{figure*}
\centering
\subfloat[]{\label{fig:sep10-ship}\includegraphics[width=0.4\textwidth]{figs/sep10-canon-cruise.png}}
\subfloat[]{\label{fig:sep10-transects}\includegraphics[width=0.40\textwidth]{figs/september-iterations.pdf}}
  \caption{\small{The offshore \can ESP-drifter following cruise
      \cite{das11b}. Fig. \ref{fig:sep10-ship} shows the ESP drifter
      foreground with the R/V \emph{Western Flyer} with the image
      taken from the R/V \emph{Zephyr}. Fig \ref{fig:sep10-transects}
      shows the 86 km transect of the Dorado following the ESP drifter
      about 100 Nm from California shores.}}
\label{fig:gulper}
\end{figure*}

These theoretic\kcomment{al} results were validated in a five-day long
off-shore deployment carried out in September $2010$, $160$ kms off
the California coast (Fig.~\ref{fig:sep10-ship}). A specialized
drifter with a genomic sensor \cite{scholin09} hanging $20$ m below,
performing in-situ identification of micro-organisms was the
target. The experiment was supported by crews on two support vessels,
the R/V \emph{Western Flyer} and the R/V \emph{Zephyr}. The
\emph{Flyer} visited the drifter every four hours to carry out a
series of ship-based sampling experiments and lab analysis on water
samples to ground-truth the drifter sensor data. The \emph{Zephyr} was
meanwhile focused on Lagrangian observation studies using the Dorado
with the goal to monitor the nutrient budget at the perimeter of a
$1$km $X 1$km water patch around the advecting drifter while the AUV
performed a transformed box pattern around the drifter. A number of
logistical issues were kept in mind while designing and executing the
experiment. Each iteration began with the latest drifter update
(position and velocity) received from the drifter through a satellite
link which was transmitted to the vehicle for in-situ adaptation. \rx
used this input to compute the $5$ waypoints necessary for an
iteration of the box pattern, with the AUV traveling at constant
velocity in the earth frame. Waypoints were computed once at the
beginning of the survey with the AUV surfacing once for every survey,
with each survey lasting $\sim1$ to $1.5$ hours. A total of 60 surveys
were attempted over the course of the $5$
days. Fig. \ref{fig:sep10-transects} shows the overall transects over
the course of the mission, with gaps in between days when the AUV was
being recharged. Each deployment lasted $\sim12$ hours. The black dots
in the figure show the beginning of each iteration when drifter
updates were received by the AUV.

\begin{figure}[t]
\centering
\includegraphics[width=0.7\textwidth]{figs/oct10-ballet.pdf}
\caption{\small An October 2010 robot ``ballet'' where a smaller
  slower AUV tracked a patch center, sending periodic updates to the
  Dorado using a setup shown in Fig. \ref{fig:dta-setup}. The Google
  Earth post-hoc image shows {\color{red}red} track lines of the \rx
  commanded Dorado with the {\color{yellow}yellow} transects of a
  separate AUV platform. {\color{yellow}Yellow} dots show the track
  line of a GPS-enabled drifter used for seeding the experiment over
  the Monterey Canyon with Santa Cruz, California to the right of the
  figure.}
\label{fig:ballet}
\end{figure}

@@

\begin{figure*}
\centering
\subfloat[]{\label{fig:odss-mtg}\includegraphics[width=0.325\textwidth]{figs/october-canon-meeting.png}}
\subfloat[]{\label{fig:odss-virtual}\includegraphics[width=0.37\textwidth]{figs/virtual-drifter-tool.png}}
\subfloat[]{\label{fig:odss-ipad}\includegraphics[width=0.325\textwidth]{figs/photo-ipad-2.jpg}}
\caption{\small Fig. \ref{fig:odss-mtg} shows a 2010 \can meeting with
  the \od in use. Fig. \ref{fig:odss-virtual} shows a drifter tracking
  tool used during \can planning; it used a physical model of advected
  surface currents from \texttt{CeNCOOS} \cite{cencoos} to plan AUV
  deployments. Fig. \ref{fig:odss-ipad} shows \od use on an iPad.}
\label{fig:odss2}
\end{figure*}

@@

Operationally, the \od has become a software system that the
scientist consults during a \can field campaign. S/he consults
up-to-date model outputs, synoptic views of ocean color from remote
sensing data and weather conditions for the next $24$-hour planning
cycle and then queries the system to compute candidate sampling
locations. In the near future, the \od will consult its situational
awareness database, incorporate model outputs projecting these outputs
into the space-time region of interest to generate candidate
hypotheses with associated utility metrics. With the scientist's
guidance the automated planner would then determine the most viable
robotic option(s) to be dispatched (underwater, surface or aerial),
given available resources, distance of travel, required payload and
the environment. For instance, unpowered autonomous gliders would be
chosen for longer-duration missions with smaller payload requirements
in areas of moderate to weak ocean currents. The automated planner
would balance these criteria to compute a viable trajectory plan for
the selected vehicle. Vehicles with more controllability would be
given a specified area to survey and the path generated \situ. Goals
are communicated to the robot after analysis on shore for selective
targeting of features by the conglomerate of vehicles in the ocean
being tracked by the \od. Thus far, the \od has provided a simple
yet functional tool for scientists to use in the daily complex
decision making and the following experiment and survey design
needs. It has helped train scientists to think of expanding their
repertoire of tools for the scientific process and in the process it
enabled a mixed team of engineers and scientists to interact at the
functional level of the systems' capabilities.
